---
title: "Interactive Visualizations for Takunda"
author: "Shangjun Yao"
date: "2025-04-23"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, message=FALSE, warning=FALSE, echo= FALSE}
library(ggplot2)
library(plotly)
library(dplyr)
library(DT)
library(tidyr)
library(flexdashboard)
```


# Dataset description
## Tab A
### Dataset Description {.tabset}

#### Overview
This project investigates how social capital influences household coping strategies (CSI) and livelihood coping strategies (LCSI) in Zimbabwe within the Takunda program over a year in February and August 2024 (Round 3 and Round 4).

#### Data Details
- **Sample size**: Round 3 (n = 483), Round 4 (n = 280)  
- **Study population**: Households in food-insecure regions of Zimbabwe  
- **Collection method**: Structured in-person surveys  
- **Confidentiality**: Dataset is private and provided for internal academic analysis only

#### Key Variables
- **Social Capital Index (SCI):** Bonding and bridging capital from internal/external help
- **Coping Strategy Index (CSI):** Short-term food insecurity coping behavior
- **Livelihood Coping Strategy Index (LCSI):** Long-term livelihood and asset coping behavior

#### Real-World Relevance
This dashboard illustrates the positive impact of social capital and resilience-building interventions on household food security. The findings can inform future program strategies for vulnerable communities in Zimbabwe and beyond.

[GitHub Repository](https://github.com/ShangjunYao/Takunda-Project)

# Interactive visualizations
## Tab B
## Social Capital vs. Coping Strategy Index {.tabset}

```{r, echo= FALSE}
Round3_updated_data <- read.csv("/Users/shangjunyao/Desktop/Capstone/Takunda data/Round3_updated_data.csv")
Round4_updated_data <- read.csv("/Users/shangjunyao/Desktop/Capstone/Takunda data/Round4_updated_data.csv")
```

```{r widget1, warning=FALSE, message=FALSE, echo= FALSE}
p1 <- ggplot(Round3_updated_data, aes(x = Social_Capital_Index, y = CSI, color = 'Round 3')) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_point(data = Round4_updated_data, aes(x = Social_Capital_Index, y = CSI, color = 'Round 4'), alpha = 0.6) +
  geom_smooth(data = Round4_updated_data, aes(x = Social_Capital_Index, y = CSI, color = 'Round 4'), method = 'lm', se = FALSE) +
  labs(title = "Social Capital vs. Coping Strategy Index (CSI)",
       x = "Social Capital Index (0–100)",
       y = "Coping Strategy Index (CSI)",
       color = "Survey Round") +
  theme_minimal()

ggplotly(p1, tooltip = c("x", "y", "color"))
```

#### Key Takeaways
- Households with higher social capital tend to use fewer short-term coping strategies.
- This pattern is visible in both survey rounds, with Round 4 showing overall lower Coping strageties scores, suggesting improved food security over time.

## Household Coping Strategies Distribution {.tabset}

```{r widget2, warning=FALSE, message=FALSE, echo= FALSE}
combined_data <- bind_rows(
  Round3_updated_data %>% mutate(Round = "Round 3"),
  Round4_updated_data %>% mutate(Round = "Round 4")
)

long_data <- combined_data %>%
  pivot_longer(cols = c(CSI, LCSI), names_to = "Coping_Strategy", values_to = "Index_Score")

boxplot_widget <- ggplot(long_data, aes(x = Coping_Strategy, y = Index_Score, fill = Round)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  labs(title = "Household Coping Strategies Distribution",
       x = "Coping Strategy Type",
       y = "Index Score",
       fill = "Survey Round") +
  theme_minimal() +
  scale_fill_manual(values = c("Round 3" = "#E69F00", "Round 4" = "#56B4E9"))


plotly::ggplotly(boxplot_widget, tooltip = c("x", "y", "fill"))
```

#### Key Takeaways
- Both Coping Strategy index and Livelihood Coping Strategy index scores were lower in Round 4, suggesting reduced use of coping strategies.
- The decrease in Livelihood Coping Strategy index was more pronounced, indicating potential long-term improvement in household resilience.


